---
title: "DATA 152 HW4 - Tyler Bontrager"
date: "Due February 22, 2023 @ 11:59pm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

** CHANGE YOUR NAME IN THE HEADER and then complete the following: **

## QUESTION 1

### 1A
Simulate 10000 standard normal random variables by running `rnorm(10000,0,1)`.  Name this variable "normSim".

```{r}
normSim = rnorm(10000,0,1)
```

### 1B
Create a histogram of your data.

```{r}
tibble(normSim) %>%
  ggplot(aes(x=normSim)) +
  geom_histogram(color="black",fill="cyan4")

```

### 1C
What percent of your observations fall 2 standard deviations below the mean?  In other words, what percent of your z scores (i.e. standardized values) are less than 2?

```{r}
nsTibble = tibble(normSim)
mean(normSim)

z = (normSim-mean(normSim))/sd(normSim)

nsTibble %>%
  mutate(
    z_score = z
  )

nsTibble %>%
  ggplot(aes(x=z,y=..density..)) +
  geom_histogram(color="black",fill="cyan4")

pnorm(-2,0,1)

```

ANSWER: 2.28% (did I do this correctly?)

### 1D
Use `pnorm` to calculate the exact probability a standard normal variable is less than -2.

```{r}
pnorm(-2,0,1) #?

```

ANSWER: 2.28%?

### 1E
Why are your answers in 1C and 1D different?

ANSWER: I couldn't seem to find the observed number of z scores 2 sds below the mean, but I'd imagine they'd be off because our data isn't supposed to be perfectly normal.

## QUESTION 2

In triathlons, it is common for racers to be placed into age and gender groups.  Friends Jed and Heather both completed the Hermosa Beach Triathlon, where Jed competed in the "Men, ages 30-34" group while Heather competed in the "Women, Ages 25-29" group.  Jed completed the race in 1:22:29 (4948 seconds), while Heather completed the race in 1:31:53 (5513 seconds).  They are curious about how they did within their respective groups.  Here's information about the performance of their groups:

* The finishing times of the "Men, ages 30-34" group has a mean of 4313 seconds with a standard deviation of 583 seconds.
* The finishing times of the "Women, Ages 25-29" group has a mean of 5261 seconds with a standard deviation of 807 seconds.  The distributions of finishing times for both groups are approximately Normal.

### 2A
What are the z-scores for Jed's and Heather's finishing times?

```{r}
jz = (4948-4313)/583
hz = (5513-5261)/807

#Jed's Z-score:
jz
#Heather's Z-score:
hz
```

ANSWER: Jed: 1.09; Heather: 0.31

### 2B
Did Jed or Heather perform better in their respective groups?  Explain your reasoning.

ANSWER: They both performed better in their respective groups. Jed performed better with respect to Heather's z-score, where he had scored over one standard deviation above the mean of his group whereas Heather performed in a better yet less significant way.

### 2C
What percent of triathletes performed better than Heather?  Remember: a better performance corresponds to a faster finish.

```{r}
1-pnorm(hz,0,1)

```

ANSWER: 37.7% of triathletes performed better than Heather.

### 2D
What percent of triathletes performed better than Jed?
Remember: a better performance corresponds to a faster finish.

```{r}
1-pnorm(jz,0,1)

```

ANSWER: 13.8% of triathletes performed better than Jed.

### 2E
Use the 68-95-99.7 Rule to calculate the range of times that encompasses 95% of all run times for the "Women, Ages 25-29" group.

```{r}

0-2*1
0+2*1

5261-2*807
5261+2*807

```

ANSWER: 95% of run times are expected to be between -2 and 2 standard deviations from the mean. This corresponds to between 3647 and 6875 seconds.

## QUESTION 3
The average daily high temperature in June in LA is 77F with a standard deviation of 5F.  Suppose that the temperatures in June closely follow a normal distribution.

```{r}
tempSim = rnorm(10000,77,5)
tibble(tempSim) %>%
  ggplot() +
  geom_histogram(aes(x=tempSim,y=..density..),color="black",fill="purple")
tempmean=77
tempsd=5

```

ANSWER: Yes?

### 3A
What is the probability of observing an 83F temperature or higher in LA during a randomly chosen day in June?

```{r}
1-pnorm(83,tempmean,tempsd)

```

ANSWER: 11.5%

### 3B
How cool are the coldest 10% of the days during June in LA?

```{r}
qnorm(.10,tempmean,tempsd)

```

ANSWER: 70.6 degF

## QUESTION 4
Cholesterol levels for women aged 20 to 34 follow an approximately normal distribution with mean 185 milligrams per deciliter (mg/dl).  Women with cholesterol levels above 220 mg/dl are considered to have high cholesterol.  About 18.5% of women fall into the "high cholesterol" category.  What is the standard deviation of cholesterol levels for women aged 20 to 34?


```{r}
cholSim = rnorm(10000,77,5)
tibble(cholSim) %>%
  ggplot() +
  geom_histogram(aes(x=cholSim,y=..density..),color="black",fill="purple")
cholmean=185
#cholstd = ?
#standardize

# z=(x-u)/o
# 18.5% is 1 standard deviation above the mean => z=1
# 1=(x-u)/o
# o=(x-u)
# o=(220-185) => o=35 mg/dL

```

ANSWER: 35 mg/dL


***
*INSTRUCTIONS TO SUBMIT* Knit this into an html document and submit BOTH the Rmd and the html file in WISE.  Name both files using this naming convention: LASTNAME_FIRSTNAME_DATA152_HW4.Rmd (or html).  For example, Dr. Gore's submissions would be named GORE_KRISTEN_DATA154_HW1.Rmd and GORE_KRISTEN_DATA152_HW4.html.  Failure to follow these instructions will result in a 10-point deduction.