---
title: "data151eda2"
author: "Tyler Bontrager"
date: "2022-10-25"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(ggplot2)
?ggplot2
```

```{r}
# IMPORTING DATASETS
tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')
tc = tuition_cost

tuition_income <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv') 
ti = tuition_income

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')
sp = salary_potential

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')
ht = historical_tuition

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv')
ds = diversity_school
```

```{r}
# Time to explore the data!
table(tc$state,tc$degree_length)
```
```{r}
prop.table(table(tc$degree_length))
```


```{r}
prop.table(table(gatheredtc$state_code,gatheredtc$degree_length))
```

**Proportions of Schools by Degree Factor and State**
```{r}
propDegByState = gatheredtc %>%
  group_by(degFactor,state_code) %>%
  summarise(n = n()) %>%
  mutate(prop = n / sum(n)) %>%
  arrange(desc(prop))
propDegByState

ggplot(propDegByState, aes(x=state_code, y=prop, color=degFactor)) + 
  geom_point() +
  ggtitle("States with the most schools (4/2 year) [MINIMALISTIC VISUALIZATION]") +
  xlab('States, in alphabetical order') +
  ylab('Proportion') +
  theme_bw()+theme(
    legend.title = element_blank())
```
Here, we can see that the x-axis is chaotic because of the fact that there are 53 discrete x-values. We see color representing the degree factor of the school: is it a two- or four-year college? A conclusion that we might be able to better draw from the table more so than the visualization is that California has a lot of schools; in fact, the most proportionally than any other state!

```{r}
table(tc$state)
```



This is a graph of the number of higher-education schools in each U.S. territory and state.

```{r}
ggplot(gatheredtc, aes(x=state_code))+
  geom_bar()
```

```{r}
#indexing each row in the table to have a unique identifier
tc$index <- 1:nrow(tc)

tc
```

```{r}
## Joint distributions
#tc2way<-tc %>%
#  group_by(state_code, degree_length)%>%
#  mutate(freq=sum(Freq))

#tc2way
```

```{r}
tc_with_count2 = tc %>%
  group_by(state_code) %>%
  mutate(school_count = n())

tc_with_count2

```


```{r}
tcFactored = tc %>%
  mutate(degFactor = as.factor(degree_length))

tcFactored
```

```{r}
ggplot(tcFactored, aes(x=state_code,fill=degFactor))+
  geom_bar(position="dodge")
```




## Numeric Summaries

```{r}
tcFactored = tc %>%
  mutate(degFactor = as.factor(degree_length))

tcFactored
```

```{r}
str(tcFactored)
```

```{r}
gatheredtc = tcFactored %>%
  filter(tc$type!='Other') %>%
  gather(key="in_out", value="totalCost",c(in_state_total,out_of_state_total))
gatheredtc
```


```{r}
ggplot(tcFactored, aes(x=in_state_tuition)) + geom_histogram()
```

```{r}
ggplot(tcFactored, aes(x=out_of_state_tuition))+geom_histogram()
```

**Total Cost of Attendance by Student Residence and College Type**

```{r}
ggplot(gatheredtc,aes(x=totalCost, color=in_out)) +
  geom_histogram() +
  expand_limits(x=80000) +
  ggtitle("Distribution of cost-of-attendance by student residence")+
  xlab("Cost of attendance ($)") +
  ylab("Count") +
  theme_bw()+theme(
    legend.title = element_blank(),
    ) + scale_fill_discrete(name = "Student Residence", labels = c("In-state", "Out-of-state"))

ggplot(gatheredtc,aes(x=totalCost, color=type)) +
  geom_histogram() +
  expand_limits(x=80000) +
  ggtitle("Distribution of cost-of-attendance by college type")+
  xlab("Cost of attendance ($)") +
  ylab("Count") +
  theme_bw()+theme(
    legend.title = element_blank())

ggplot(gatheredtc,aes(x=totalCost, color=in_out)) +
  geom_density() +
  expand_limits(x=80000) +
  ggtitle("Distribution of cost-of-attendance by student residence")+
  xlab("Cost of attendance ($)") +
  ylab("Density") +
  theme_bw()+theme(
    legend.title = element_blank(),
    ) + scale_fill_discrete(name = "Student Residence", labels = c("In-state", "Out-of-state"))

ggplot(gatheredtc,aes(x=totalCost, color=type)) +
  geom_density() +
  expand_limits(x=80000) +
  ggtitle("Distribution of cost-of-attendance by college type")+
  xlab("Cost of attendance ($)") +
  ylab("Density") +
  theme_bw()+theme(
    legend.title = element_blank())


# ggplot(tcFactored, aes(x=in_state_total))+geom_histogram()+expand_limits(x=80000,y=430) +
#   ggtitle("Total cost of attendance for in-state students")+ # for the main title
#   xlab("Cost of attendance")+ # for the x axis label
#   ylab("Count") # for the y axis label
# 
# 
# ggplot(tcFactored, aes(x=out_of_state_total))+geom_histogram()+expand_limits(x=80000,y=430) +
#   ggtitle("Total cost of attendance for out-of-state students")+ # for the main title
#   xlab("Cost of attendance")+ # for the x axis label
#   ylab("Count") # for the y axis label


#ggtitle(label) # for the main title
#xlab(label) # for the x axis label
#ylab(label) # for the y axis label
#labs(...) # for the main title, axis labels and legend titles

```



















