---
title: "Regression"
author: "Tyler Bontrager, Ganesh Singh"
date: "2022-11-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(ggplot2)
?ggplot2
```

```{r}
# IMPORTING DATASETS
tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')
tc = tuition_cost

tuition_income <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv') 
ti = tuition_income

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')
sp = salary_potential

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')
ht = historical_tuition

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv')
ds = diversity_school
```

```{r}
str(tc)
str(ti)
str(sp)
str(ht)
str(ds)
```



```{r}
tcFactored = tc %>%
  mutate(degFactor = as.factor(degree_length))
tcFactored

tcFacJoinSp = tcFactored %>%
  inner_join(sp, by=c("name"="name"))
tcFacJoinSp

```
```{r}
tcFactored
sp
tcFacJoinSp
```


```{r}
ggplot(sp, aes(stem_percent,early_career_pay)) + geom_point() +
  geom_smooth(method="lm",se=FALSE)+
  ggtitle("Early career pay against prevalence of STEM majors in a school")+
  xlab("Prevalence of STEM majors (%)")+
  ylab("Early-career pay ($)")
```
```{r}
ECSPmodel = lm(early_career_pay~stem_percent,data=tcFacJoinSp)
summary(ECSPmodel)
```
Slope: 323.30
Y-intercept: 45584.50


```{r}
cor(x=sp$stem_percent,y=sp$early_career_pay)
cor(x=sp$stem_percent,y=sp$mid_career_pay)
```

Moderately strong positive association between prevalence of STEM majors at a school and general alumni early-career earnings.


Boring. STEM jobs tend to pay really well. Let's do something fun. Hear me out...


```{r}
ggplot(tcFacJoinSp, aes(x=stem_percent,y=in_state_tuition,color=type)) + 
  geom_point() +
  geom_smooth(method="lm",se=FALSE,lty=2) +
  ggtitle("Cost of tuition for in-state students against prevalence of STEM majors")+
  xlab("Prevalence of STEM majors (%)")+
  ylab("In-state tuition cost ($)")
```
(This graph doesn't show what was seen first for your convenience, but I will be sure to remove the color for the presentation)

```{r}
ggplot(tcFacJoinSp, aes(x=stem_percent,y=out_of_state_tuition,color=type)) + 
  geom_point() +
  geom_smooth(method="lm",se=FALSE,lty=2) +
  ggtitle("Cost of tuition for out-state students against prevalence of STEM majors")+
  xlab("Prevalence of STEM majors (%)")+
  ylab("Out-state tuition cost ($)")
```


```{r}
ISSPmodel=lm(in_state_tuition~stem_percent,data=tcFacJoinSp)
OSSPmodel=lm(out_of_state_tuition~stem_percent,data=tcFacJoinSp)

summary(ISSPmodel)
summary(OSSPmodel)
```
In-state
Slope: 258.53
Y-intercept: 21958.94

Out-of-state
Slope: 289.00
Y-intercept: 25835.03

Notes for story-telling:

Hypothesis: stem_percent high => stem equipment high; cost++?

room and board not considered because r&b charged by a non-academic department, and stem is academic
in-state plot revealed cluster
out-of-state students do not generally enjoy benefits of tuition discounts at public universities

Hypothesis: public => less burden on student due to public/instate status => generally less tuition, but still guessing stem equipment drives up cost.

```{r}
cor(tcFacJoinSp$in_state_total,tcFacJoinSp$stem_percent)

#duh, cor doesn't care about the clusters.
```
```{r}
tcFacJoinSpPUB = tcFacJoinSp %>%
  filter(type=="Public")
tcFacJoinSpPUB

tcFacJoinSpPRIV = tcFacJoinSp %>%
  filter(type=="Private")
tcFacJoinSpPRIV

```

```{r}
cor(tcFacJoinSpPRIV$in_state_tuition,tcFacJoinSpPRIV$stem_percent)
cor(tcFacJoinSpPUB$in_state_tuition,tcFacJoinSpPUB$stem_percent)
```

```{r}
cor(tcFacJoinSpPRIV$out_of_state_tuition,tcFacJoinSpPRIV$stem_percent)
cor(tcFacJoinSpPUB$out_of_state_tuition,tcFacJoinSpPUB$stem_percent)
```

Moderately weak positive association between tuition cost and prevalence of stem majors at a school. But only MODERATELY! There still is some good correlation here. Other ideas?


Observation notes:

Correlations for both seem to be generally higher than that of the one taken before splitting.

Question: what is a stronger explanation for expensive schools? Are there other factors beyond funding STEM departments?

This dataset does not contain readily-usable data on ivy-league status or reputation. What if schools use their alumni's early career salary as a selling point to justify higher prices?















