---
title: 'DATA151: Comparing Distributions'
author: "Kitada Smalley"
output:
  html_document:
    collapsed: no
    toc: yes
    toc_depth: 5
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
---

### Learning Objectives

In this lesson students will compare distributions from multiple populations of interest using: 

* `dplyr`: `group_by` and `summarise`
* side-by-side `boxplots`
* side-by-side `violin` plots
* `beeswarm`plots
* faceting


#### Step 0: Library Tidyverse

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

#### Step 1: Load the Data

```{r}
aqi<-read.csv("https://raw.githubusercontent.com/kitadasmalley/DATA151/main/Data/fireAQI_OrCoWa_10192022.csv", 
                   header=TRUE)

```

#### Step 2: Numeric Summaries

How does the air quality compare in Oregon, Washington, and Colorado?

```{r}
## LET'S START WITH MEAN and STANDARD DEV
aqi%>%
  group_by(State)%>%
  summarise(n=n(), 
            avgAQI=mean(AQI,na.rm=TRUE), 
            sdAQI=sd(AQI,na.rm=TRUE), 
            NAs=sum(is.na(AQI)))
```


Now let's add more metrics!  Let's use the quartiles.  Cutting our data into quartiles means that we have split the data into four even parts.  

* $Q_1$: The first quartile is the 25th percentile 
* $Q_2$ (Median): The second quartile is also known as the median and is the 50th percentile
*  $Q_3$: The third quartile is the 75th percentile 

We can use the `quantile()` function to get any desired quantile from our data set.  Quantile is synonymous with percentile.  When we use the quantile function we specify the fraction of data below a desired cut off point. 

```{r}
# The first quartile
quantile(aqi$AQI, prob=.25, na.rm = TRUE)
```


We can do this for each state: 

```{r}
aqi%>%
  group_by(State)%>%
  summarise(n=n(), 
            min=min(AQI, na.rm=TRUE),
            Q1=quantile(AQI,prob=0.25,na.rm=TRUE),
            med=median(AQI,na.rm=TRUE), #or quantile(AQI,prob=0.5,na.rm=TRUE)
            Q3=quantile(AQI,prob=0.75,na.rm=TRUE),
            max=max(AQI,na.rm=TRUE))
```

What do you observe?

#### Step 3: Side-by-side Boxplots

```{r}
# BOXPLOT
ggplot(aqi,aes(x=State,y=AQI,fill=State))+geom_boxplot()
```

#### Step 4: Side-by-side Violin Plots

```{r}
# VIOLIN

##INSERT CODE HERE ##
```

#### Step 5: Dot Plots

You could make a dot plot!
```{r}
# Dot plot

##INSERT CODE HERE ##
```

#### Step 6: Beeswarm Plots

```{r}
#install.packages('ggbeeswarm')
library(ggbeeswarm)

##INSERT CODE HERE ##
```

#### Step 7: Faceting

Faceting can be used with any type of plot to separate subgroups.

##### Faceting Histograms

Here is a histogram where the y-axis is count.

```{r}
### Histogram (Counts)
##INSERT CODE HERE ##
```

We can change the y-axis to density (proportions).

```{r}
### Histogram (Density)
##INSERT CODE HERE ##
```

Since there is a spike in the Colorado distribution, this distorts the scale for the distributions for Oregon and Washington.  We can fix this by allowing the y-axis to freely vary. 

```{r}
### Histogram (Density - Free_y)
##INSERT CODE HERE ##
```

##### Faceting Density Plots

We can also make a density plot!

```{r}
### Density Plot (free_y)
##INSERT CODE HERE ##
```