---
title: "DATA 152 HW1 - Tyler Bontrager"
date: "Due February 1, 2023 @ 11:59pm"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(skimr)
```

**INSTRUCTIONS:** Change your name in the title section and then complete the following questions.

## QUESTION 1 (30)
In a public health study on the effects of consumption of fruits and vegetables on psychological well-being in young adults, participants were randomly assigned to 3 groups: 1) diet-as-usual 2) an ecological intervention involving text message reminders to increase their fruits and vegetable consumption + a voucher to purchase them or 3) a fruit and vegetable intervention in which participants were given 2 additional daily servings of fresh fruits and vegetables to consume on top of their normal diet.  Participants were asked to take a nightly survey on their smartphones.  Participants were student volunteers at the University of Otago, New Zealand.  At the end of the 14-day study, only participants in the 3rd group showed improvements to their psychological well-being across the 14 days, relative to the other groups.

### 1A (5)
What type of study is this?  (Experiment or Observational Study)

*ANSWER:* Experimental study

### 1B (10)
What are the explanatory and response variables?  (Explanatory=covariates or factors)

*ANSWER:* Explanatory: diet interventions. Response: psychological well-being across 14 days.

### 1C (5)
Comment on whether the results of the study can be generalized to the population.

*ANSWER:* Potentially. Humans are complex, but function similarly biologically. Depending on the size of the study, it could be a well-supported argument for some generalized claim.

### 1D (5)
Comment on whether the results of the study can be used to establish causal relationships

*ANSWER:* I think this study doesn't take into account various factors like the fact they're at a university which ostensibly strains students' mental health regularly. There seems to be a lot of confounding variables, which means that causal relationships may not be present between the fruits and vegetables and psychological well-being.

### 1E (5)
A newspaper article reporting on the study states, "The results of this study provide proof that giving young people fresh fruits and vegetables to eat can have psychological benefits, even over a brief period of time."  How would you suggest revising the statement so that it can be supported by the study?

*ANSWER:* "The results of this study show that making fresh fruits and vegetables available to eat may improve psychological benefits, even over a brief period of time."


## QUESTION 2 (20)
A city council has requested a household survey be conducted in a suburban area of their city.  The area is broken into many distinct and unique neighborhoods, some including large homes, some with only apartments, and others with a diverse mixture of housing structures.  

*INSTRUCTIONS:* For each part below, identify the sampling method (simple random sample, stratified sampling, cluster sampling, multistage) described, and **describe the statistical pros and cons of the method in the city's context**.

### 2A (5)
Randomly sample 200 households from the city

*ANSWER:* simple random sampling

### 2B (5)
Divide the city into 20 neighborhoods, and sample 10 households from each neighborhood.

*ANSWER:* stratified random sampling

### 2C (5)
Divide the city into 20 neighborhoods, randomly sample 3 neighborhoods, and then sample all households from those 3 neighborhoods.

*ANSWER:* cluster sampling

### 2D (5)
Divide the city into 20 neighborhoods, randomly sample 8 neighborhoods, and then randomly sample 50 households from those neighborhoods

*ANSWER:* multistage sampling


## QUESTION 3 (20)
An experiment showed that subjects fed the DASH diet were able to lower their blood pressure by an average of 6.7 points compared to a group fed a "control diet."  All meals were prepared by dietitians.

### 3A (5)
Why were the subjects randomly assigned to the diets instead of letting people pick what they wanted to eat?

*ANSWER:* Because humans are complex, and this should help reduce the impact of biological confounding variables.

### 3B (5)
Why were the meals prepared by dietitians?

*ANSWER:* Experts know what should be in people's diet generally speaking, and they probably know how to manipulate the variables in the experiment much more easily.

### 3C (5)
Why did the researchers need the control group?  If the DASH diet group's blood pressure was lower at the end of the experiment than at the beginning, wouldn't that prove the effectiveness of that diet?

*ANSWER:* No because it wouldn't establish a causal relationship between the DASH diet and blood pressure.

### 3D (5)
What additional information would you want to know in order to decide whether an average reduction in blood pressure of 6.7 points was statistically significant?

*ANSWER:* The change in blood pressure of the control group.

## QUESTION 4 (25)
Is there a universally "best" temperature for citrus plants grown in a greenhouse?  Suppose you have 18 citrus plants that are the same age (6 orange, 6 lemons, 6 limes).  You also opt to use a combination of natural and artificial sunlight to control for that factor. 

### 4A (10)
Create a design matrix which allows you to test 3 temperature levels (you can name them "T_low", "T_med", and "T_high" for now) for all 3 citrus species.

```{r}
# your code below
plants=c("orange","lemon","lime")
treatment=c("T_low","T_med","T_high")
designMatrix = expand.grid(plants,treatment)
designMatrix
```

### 4B (5)
Assuming a balanced design (i.e. equal # of replicates for each split), how many replicates can you run per split with 18 citrus plants?

*ANSWER:* 3

### 4C (10)
Create a full experimental design with the replicates built into the design table. 
```{r}
# your code below
designMatrix %>%
  group_by(Var1,Var2) %>%
  expand(count=1:3)

```

## QUESTION 5 (5)
What are the elements of a good experimental design?

*ANSWER:* A good experiment controls for confounding variables by randomization and blocking, strong replication potential, and establishes an argument for causality by keeping a control group.



***
*INSTRUCTIONS TO SUBMIT* Knit this into an html document and submit BOTH the Rmd and the html file in WISE.  Name both files using this naming convention: LASTNAME_FIRSTNAME_DATA152_HW1.Rmd (or html).  For example, Dr. Gore's submissions would be named GORE_KRISTEN_DATA152_HW1.Rmd and GORE_KRISTEN_DATA152_HW1.html.  Failure to follow these instructions will result in a 10-point deduction.